---
title: "Simulation"
author: "Bowei Kang; Zikai Li"
date: "2023-05-20"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dfcrm)
library(ggplot2)
library(tidyverse)
library(HDInterval)
```

# simulation study based on Figure 6

```{r}
dose = c(0.5, 1, 3, 5, 6)
p_true_1 = c(0.25, 0.30, 0.50, 0.60, 0.70)
p_true_2 = c(0.01, 0.05, 0.20, 0.30, 0.50)
N = 36
n_per_cohort = 3
ncohort = N/n_per_cohort
target = 0.33
```

# CRM 

Start from the lowest level (level 1). No skipping. 
In each batch, if prop.tox <= 1/6, upgrade; if 1/6 < prop.tox <= 1/3, keep; if prop.tox > 1/3, downgrade

## functions

```{r}
determine_next_dose = function(level_current, prop_tox_cur_batch, level_max) {
  level_new = 1
  if (level_current == 1) {
    if (prop_tox_cur_batch <=1/6) level_new = 2 
  }
  if (level_current == level_max) {
    if (prop_tox_cur_batch <=1/3) {
      level_new = level_max }
      else level_new = level_current - 1
  }
  if (level_current %in% 2:(level_max-1)) {
    if (prop_tox_cur_batch <=1/6) level_new = level_current + 1
    if (prop_tox_cur_batch > 1/6 & prop_tox_cur_batch <=1/3) {
      level_new = level_current
    }
    if (prop_tox_cur_batch > 1/3) level_new = level_current - 1
  } 
  return(level_new)
}


single_run = function(N, n_per_cohort, target, p_true, p.tox0, dose, plt = F) {
  ncohort = N/n_per_cohort
  path <- data.frame(patient = 1:N, dose = NA, tox = NA, batches = NA)

  for (i in 1:ncohort) {
    if (i == 1) {
      level_current = 1
      batch_current = 1
    }
    
    path[((i-1)*3+1):(i*3), 'dose']  = level_current
    path[((i-1)*3+1):(i*3), 'batches'] = batch_current
    path[((i-1)*3+1):(i*3), 'tox'] = rbinom(n_per_cohort, 1, p_true[level_current])

    ids_cur_batch = which(path[, 'batches'] == batch_current)
    prop_tox_cur_batch = mean(path[ids_cur_batch, 'tox'])
    level_update = determine_next_dose(level_current, prop_tox_cur_batch, 
                                       length(p_true))
    
    if (level_update != level_current) {
      batch_current = batch_current + 1
      level_current = level_update}
} # end for
  
# bayesian estimate using logistic model
# est = crm(p.tox0, target, path[,'y'], path[, 'levels'], model = 'logistic2')
# mtd = dose[which.min(abs(est$ptox - target))]
# if (plt) pp = plt_func(path, est, target, mtd, p_true, dose) else pp = NULL
# return(list(path = path, est = est, target = target, mtd = mtd, p_true = p_true, 
#             dose = dose))


# assume normal priors for params. mu = 0, sd = 1 for both. independent
est <- bcrm(stop=list(nmax=18), data=path, sdose = dose,
            ff="logit2", prior.alpha=list(4, c(0, 0), rbind(c(1, 0), c(0, 1))), 
            target.tox=target, constrain=F, sdose.calculate="median", 
            pointest="mean")

mcmc_post_params <- Posterior.rjags(tox = est$tox, notox = est$notox, 
                                sdose = est$sdose, ff = "logit2", 
                                prior.alpha = est$prior.alpha, 
                                burnin.itr = 2000, production.itr = 2000)

alpha_post_mean<- mean(mcmc_post_params[,1])
beta_post_mean <- mean(mcmc_post_params[,2])
tox_post <- (exp(log(alpha_post_mean)+beta_post_mean*est$sdose)/
               (1+exp(log(alpha_post_mean)+beta_post_mean*est$sdose)))
selector <- which.min(abs(tox_post - target))
mtd <- dose[selector]

lb <- rep(0, 5)
ub <- rep(0, 5)

for (i in 1:length(est$sdose)){
  s <-  est$sdose[i]
  tox_post_all <- (exp(log(mcmc_post_params[,1])+mcmc_post_params[,2]*s)/
               (1+exp(log(mcmc_post_params[,1])+mcmc_post_params[,2]*s)))
  hdi <- hdi(tox_post_all, prob = 0.95)
  
  lb[i] = hdi[1]
  ub[i] = hdi[2]

}



return(list(mtd = mtd, path = path, target = target, p_true = p_true, dose = dose, 
            lb = lb, ub = ub, tox_post = tox_post))
  
}

plt_func = function(path, est, target, mtd, p_true, dose) {
p1 = ggplot(data.frame(path), aes(patient, dose, col = as.factor(tox))) + geom_point() + theme_bw() + ylim(0,length(dose) + 1) 
tbl = data.frame(x = dose, p = est$ptox, lb = est$ptoxL, ub = est$ptoxU, ptrue = p_true)
p2 = ggplot(tbl) + geom_line(aes(x, p)) + geom_point(aes(x, p), shape = 21, colour = "black", fill = "white", size = 2, stroke = 1) 
p2 = p2 + geom_line(aes(x, ub), linetype = 2) + geom_point(aes(x, ub), shape = 21, colour = "black", fill = "white", size = 2, stroke = 1) + geom_line(aes(x, lb), linetype = 2) + geom_point(aes(x, lb), shape = 21, colour = "black", fill = "white", size = 2, stroke = 1) 
p2 = p2 + geom_line(aes(x, ptrue), col = 'blue') + geom_point(aes(x, ptrue), shape = 22, colour = "blue", fill = "white", size = 2, stroke = 1) 
p2 = p2 + geom_hline(yintercept = target, col = 'red', linetype = 2, linewidth = 0.8) 
p2 = p2 + geom_vline(xintercept = mtd, col = 'green4', linetype = 3, linewidth = 1)
p2 = p2 + scale_x_continuous(breaks = dose) + labs(title = '', x = 'Dose', y = 'Probability of DLT')
p2 = p2 + theme_bw() 
return(p1 + p2)
}
```

## test a single trial

```{r}
set.seed(1)
res = single_run(N = 36, n_per_cohort = 3, target = 0.33, p_true = p_true_1, p.tox0 = p_true_1, dose = dose, plt = F)
res
```

## multiple runs - case 1

```{r}
set.seed(1)
niter = 1000
res1 = replicate(niter, single_run_3(N = 36, n_per_cohort = 3, target = 0.33, p_true = p_true_1, p.tox0 = p_true_1, dose = dose, plt = FALSE))
# visit frequency
vf1 = sapply(res1[1,], function(t) {
summary = NULL
for (i in 1:length(dose)) summary = c(summary, sum(t[, 'dose'] == i))
return(summary)
})
rowSums(vf1) / (niter * N)
# mtd
table(unlist(res1[3,])) / niter
p_true_1
```

57% visits level 1 (dose = 0.5), 55% sample trials recommends mtd at level 2 (dose = 1)

## multiple runs - case 2

```{r}
set.seed(1)
niter = 1000
res2 = replicate(niter, single_run_3(N = 36, n_per_cohort = 3, target = 0.33, p_true = p_true_2, p.tox0 = p_true_1, dose = dose, plt = FALSE))
# visit frequency
vf2 = sapply(res2[1,], function(t) {
summary = NULL
for (i in 1:length(dose)) summary = c(summary, sum(t[, 'dose'] == i))
return(summary)
})
rowSums(vf2) / (niter * N)
# mtd
table(unlist(res2[3,])) / niter
p_true_2
```

31% and 33% visits level 3, 4 (dose = 3, 5), 62% sample trials recommends mtd at level 4 (dose = 5)

## visit frequency

Visit frequency of case 2 is more spread-out, which leads to a better performance.

```{r}
df_bar = data.frame(level = rep(1:length(dose), 2), prop = c(rowSums(vf1), rowSums(vf2)) / (niter * N), case = rep(c('case1', 'case2'), c(length(dose), length(dose))))
ggplot(df_bar, aes(x = level, y = prop, fill = case)) + geom_bar(stat = "identity", position = position_dodge()) + theme_bw()
```


# Improve CRM by adding extra lower levels

Add extra levels (level 1/4, 1/3, 1/2) lower than the suggested lowest level (level 1, dose = 0.5)
Start from new lowest level (level 1/4). No skipping. 
In each batch, if prop.tox <= 1/6, upgrade; if 1/6 < prop.tox <= 1/3, keep; if prop.tox > 1/3, downgrade.

## test a single trial

```{r}
dose_extd = c(0.5/4, 0.5/3, 0.5/2, 0.5, 1, 3, 5, 6)
p_true_1_extd = c(0.1, 0.15, 0.2, 0.25, 0.3, 0.5, 0.6, 0.7)
p_true_2_extd = c(0.001, 0.004, 0.007, 0.01, 0.05, 0.2, 0.3, 0.5)
set.seed(1)
res = single_run_3(N = 36, n_per_cohort = 3, target = 0.33, p_true = p_true_1_extd, p.tox0 = p_true_1_extd, dose = dose_extd, plt = T)
res
```

## multiple runs - case 1

```{r}
set.seed(1)
niter = 1000
res1 = replicate(niter, single_run_3(N = 36, n_per_cohort = 3, target = 0.33, p_true = p_true_1_extd, p.tox0 = p_true_1_extd, dose = dose_extd, plt = FALSE))
# visit frequency
vf1 = sapply(res1[1,], function(t) {
summary = NULL
for (i in 1:length(dose_extd)) summary = c(summary, sum(t[, 'dose'] == i))
return(summary)
})
rowSums(vf1) / (niter * N)
# mtd
table(unlist(res1[3,])) / niter
p_true_1_extd
```

About 20% visits level 2,3,4 (dose = 0.5/3, 0.5/2, 0.5), most (56%) sample trails recommends mtd at level 4 (dose = 1)

## multiple runs - case 2

```{r}
set.seed(1)
niter = 1000
res2 = replicate(niter, single_run_3(N = 36, n_per_cohort = 3, target = 0.33, p_true = p_true_2_extd, p.tox0 = p_true_2_extd, dose = dose_extd, plt = FALSE))
# visit frequency
vf2 = sapply(res2[1,], function(t) {
summary = NULL
for (i in 1:length(dose_extd)) summary = c(summary, sum(t[, 'dose'] == i))
return(summary)
})
rowSums(vf2) / (niter * N)
# mtd
table(unlist(res2[3,])) / niter
p_true_2_extd
```

most visits level 6, 7 (dose = 3, 5), 57% sample trials recommends mtd at level 7 (dose = 5)

## visit frequency

Both spread out. The performance difference is mitigated. 56% vs 57%. Standard CRM - 55% vs 62%.

```{r}
df_bar = data.frame(level = rep(1:length(dose_extd), 2), prop = c(rowSums(vf1), rowSums(vf2)) / (niter * N), case = rep(c('case1', 'case2'), c(length(dose_extd), length(dose_extd))))
ggplot(df_bar, aes(x = level, y = prop, fill = case)) + geom_bar(stat = "identity", position = position_dodge()) + theme_bw()
```
