---
title: "Simulation"
author: "Bowei Kang; Zikai Li"
date: "2023-05-20"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = F, message=F}
# library(dfcrm)
library(bcrm)
library(ggplot2)
library(tidyverse)
library(HDInterval)
library(ggpubr)
```

# simulation study based on Figure 6

```{r}
dose = c(0.5, 1, 3, 5, 6)
p.tox0 = c(0.05, 0.10, 0.15, 0.33,0.50) 
p_true_1 = c(0.25, 0.30, 0.50, 0.60, 0.70)
p_true_2 = c(0.01, 0.05, 0.20, 0.30, 0.50)
N = 36
n_per_cohort = 3
ncohort = N/n_per_cohort
target = 0.33
```

# CRM 

Start from the lowest level (level 1). dose skipping is allowed. Use the recommended level by bcrm as the next dose.

## functions

```{r}
single_run = function(N, n_per_cohort, target, p_true, p.tox0, dose, plt = F) {
ncohort = N/n_per_cohort
path = data.frame(patient = 1:N, dose = NA, tox = NA, batches = NA)
# Here dose represent dose levels of recruited patients, ranging from 1, ..., k
# generate a path
for (i in 1:ncohort) {
if (i == 1) {
level_current = 1
batch_current = 1
}
path[((i-1)*3+1):(i*3), 'dose']  = level_current
path[((i-1)*3+1):(i*3), 'batches'] = batch_current
path[((i-1)*3+1):(i*3), 'tox'] = rbinom(n_per_cohort, 1, p_true[level_current])
# estimate p.tox profile based on current path
ids_batch_sofar = which(path[, 'batches'] %in% 1:batch_current)
level_update_bcrm = bcrm(stop = list(nmax = length(ids_batch_sofar)), data = path[ids_batch_sofar, ], dose = dose, p.tox0 = p.tox0, ff="logit2", prior.alpha=list(4, c(1, 1), rbind(c(1, 0), c(0, 1))), target.tox = target, constrain = F, pointest = "mean", method = 'rjags')
# pointest: Which summary estimate of the posterior distribution should be used to choose the next dose.
level_update = as.numeric(level_update_bcrm$ndose[[1]]$ndose)
# update current level and batch 
if (level_update != level_current) {
batch_current = batch_current + 1
level_current = level_update
} 
} # end for
  
# assume normal priors for params. mu = 0, sd = 1 for both. independent
final_est = bcrm(stop = list(nmax = N), data = path, dose = dose, p.tox0 = p.tox0, ff="logit2", prior.alpha = list(4, c(1, 1), rbind(c(1, 0), c(0, 1))), target.tox = target, constrain = F, pointest="mean", method = 'rjags')
ptox = final_est$ndose[[1]]$est
ptoxL = final_est$ndose[[1]]$quantiles[1,]
ptoxU = final_est$ndose[[1]]$quantiles[5,]
mtd = which.min(abs(ptox - target))
if (plt) pp = plt_func(path, ptox, ptoxL, ptoxU, target, mtd, p_true, dose) else pp = NULL
return(list(path = path, ptox = ptox, ptoxL = ptoxL, ptoxU = ptoxU, mtd = mtd, plt = pp))
}


# plot function

plt_func = function(path, ptox, ptoxL, ptoxU, target, mtd, p_true, dose) {
p1 = ggplot(data.frame(path), aes(patient, dose, col = as.factor(tox))) + geom_point() + theme_bw() + ylim(0, length(dose) + 1) 
tbl = data.frame(x = dose, p = ptox, lb = ptoxL, ub = ptoxU, ptrue = p_true)
p2 = ggplot(tbl) + geom_line(aes(x, p)) + geom_point(aes(x, p), shape = 21, colour = "black", fill = "white", size = 2, stroke = 1) 
p2 = p2 + geom_line(aes(x, ub), linetype = 2) + geom_point(aes(x, ub), shape = 21, colour = "black", fill = "white", size = 2, stroke = 1) + geom_line(aes(x, lb), linetype = 2) + geom_point(aes(x, lb), shape = 21, colour = "black", fill = "white", size = 2, stroke = 1) 
p2 = p2 + geom_line(aes(x, ptrue), col = 'blue') + geom_point(aes(x, ptrue), shape = 22, colour = "blue", fill = "white", size = 2, stroke = 1) 
p2 = p2 + geom_hline(yintercept = target, col = 'red', linetype = 2, linewidth = 0.8) 
p2 = p2 + geom_vline(xintercept = dose[mtd], col = 'green4', linetype = 3, linewidth = 1)
p2 = p2 + scale_x_continuous(breaks = dose) + labs(title = '', x = 'Dose', y = 'Probability of DLT')
p2 = p2 + theme_bw() 
return(ggarrange(p1, p2))
}

```

## test a single trial

```{r}
set.seed(1)
res = single_run(N = 36, n_per_cohort = 3, target = 0.33, p_true = p_true_1, p.tox0 = p.tox0, dose = dose, plt = T)
res
```

## multiple runs - case 1

```{r}
set.seed(1)
niter = 500
res1 = replicate(niter, single_run(N = 36, n_per_cohort = 3, target = 0.33, p_true = p_true_1, p.tox0 = p.tox0, dose = dose, plt = FALSE))
# visit frequency
print("Percent of patients assigned to each dose over all trials")
vf1 = sapply(res1[1,], function(t) {
summary = NULL
for (i in 1:length(dose)) summary = c(summary, sum(t[, 'dose'] == i))
return(summary)
})
rowSums(vf1) / (niter * N)
# mtd
print("Proportion of simulated trials recommending dose as MTD")
# table(unlist(res1[5,])) / niter
sapply(1:length(dose), function(t) mean(unlist(res1[5,]) == t))

# True ptox
print("True p's")
p_true_1
```


```{r}
# [1] "Percent of patients assigned to each dose over all trials"
# [1] 0.44900000 0.23800000 0.20750000 0.05933333 0.04616667
# [1] "Proportion of simulated trials recommending dose as MTD"
# [1] 0.374 0.390 0.218 0.018 0.000
# [1] "True p's"
# [1] 0.25 0.30 0.50 0.60 0.70
```


## multiple runs - case 2

```{r}
set.seed(1)
niter = 500
res2 = replicate(niter, single_run(N = 36, n_per_cohort = 3, target = 0.33, p_true = p_true_2, p.tox0 = p.tox0, dose = dose, plt = FALSE))
# visit frequency
print("Percent of patients assigned to each dose over all trials")
vf2 = sapply(res2[1,], function(t) {
summary = NULL
for (i in 1:length(dose)) summary = c(summary, sum(t[, 'dose'] == i))
return(summary)
})
rowSums(vf2) / (niter * N)
# mtd
print("Proportion of simulated trials recommending dose as MTD")
# table(unlist(res2[5,])) / niter
sapply(1:length(dose), function(t) mean(unlist(res2[5,]) == t))

# True ptox
print("True p's")
p_true_2
```


```{r}
# [1] "Percent of patients assigned to each dose over all trials"
# [1] 0.08816667 0.00950000 0.11066667 0.45316667 0.33850000
# [1] "Proportion of simulated trials recommending dose as MTD"
# [1] 0.00 0.00 0.11 0.67 0.22
# [1] "True p's"
# [1] 0.01 0.05 0.20 0.30 0.50
```


# Bonus

At the 1/3 point of the trail (after 12th patient), evaluate the dose visit distribution. If the distribution is too skewed (level 1 attains highest freq), an extra lower level 0 will be added. Continue the previous trial design otherwise.

```{r}
dose = c(0.5, 1, 3, 5, 6)
p.tox0 = c(0.05, 0.10, 0.15, 0.33,0.50) 
p_true_1 = c(0.25, 0.30, 0.50, 0.60, 0.70)
p_true_2 = c(0.01, 0.05, 0.20, 0.30, 0.50)
dose_add = c(0.25, dose)
p.tox0_add = c(0.025, p.tox0)
p_true_1_add = c(0.125, p_true_1)
p_true_2_add = c(0.005, p_true_2)
N = 36
n_per_cohort = 3
ncohort = N/n_per_cohort
target = 0.33
train_frac = 1/3
```

## functions 

```{r}
single_run_add_dose = function(N, n_per_cohort, target, p_true_add, p.tox0_add, dose_add, train_frac = 1/3, plt = F) {
p_true = p_true_add[-1]
p.tox0 = p.tox0_add[-1]
dose = dose_add[-1]
ncohort = N/n_per_cohort
path = data.frame(patient = 1:N, dose = NA, tox = NA, batches = NA)
# Here dose represent dose levels of recruited patients, ranging from 1, ..., k
# generate the path for the first half
for (i in 1:(ncohort * train_frac)) {
if (i == 1) {
level_current = 1
batch_current = 1
}
path[((i-1)*3+1):(i*3), 'dose']  = level_current
path[((i-1)*3+1):(i*3), 'batches'] = batch_current
path[((i-1)*3+1):(i*3), 'tox'] = rbinom(n_per_cohort, 1, p_true[level_current])
# estimate p.tox profile based on current path
ids_batch_sofar = which(path[, 'batches'] %in% 1:batch_current)
level_update_bcrm = bcrm(stop = list(nmax = length(ids_batch_sofar)), data = path[ids_batch_sofar, ], dose = dose, p.tox0 = p.tox0, ff="logit2", prior.alpha=list(4, c(1, 1), rbind(c(1, 0), c(0, 1))), target.tox = target, constrain = F, pointest = "mean", method = 'rjags')
# pointest: Which summary estimate of the posterior distribution should be used to choose the next dose.
level_update = as.numeric(level_update_bcrm$ndose[[1]]$ndose)
# update current level and batch 
if (level_update != level_current) {
batch_current = batch_current + 1
level_current = level_update
} 
} # end for

# 1/3 way evaluation
# shift one level
path[1:(N*train_frac), 'dose'] = path[1:(N*train_frac), 'dose'] + 1
tt = table(path[1:(N*train_frac), 'dose'])
add_flag = F
# if the largest freq is attained at dose level 1
if (sum(path[1:(N*train_frac), 'dose'] == 2) == max(tt)) {
add_flag = T
level_current = 1
batch_current = batch_current + 1
} 

# second half
for (i in (ncohort*train_frac+1):ncohort) {
path[((i-1)*3+1):(i*3), 'dose']  = level_current
path[((i-1)*3+1):(i*3), 'batches'] = batch_current
path[((i-1)*3+1):(i*3), 'tox'] = rbinom(n_per_cohort, 1, p_true_add[level_current])
# estimate p.tox profile based on current path
ids_batch_sofar = which(path[, 'batches'] %in% 1:batch_current)
level_update_bcrm = bcrm(stop = list(nmax = length(ids_batch_sofar)), data = path[ids_batch_sofar, ], dose = dose_add, p.tox0 = p.tox0_add, ff="logit2", prior.alpha=list(4, c(0, 0), rbind(c(1, 0), c(0, 1))), target.tox = target, constrain = F, pointest = "mean", method = 'rjags')
# pointest: Which summary estimate of the posterior distribution should be used to choose the next dose.
level_update = as.numeric(level_update_bcrm$ndose[[1]]$ndose)
# update current level and batch 
if (level_update != level_current) {
batch_current = batch_current + 1
level_current = level_update
} 
} # end for

# assume normal priors for params. mu = 0, sd = 1 for both. independent
final_est = bcrm(stop = list(nmax = N), data = path, dose = dose_add, p.tox0 = p.tox0_add, ff="logit2", prior.alpha = list(4, c(1, 1), rbind(c(1, 0), c(0, 1))), target.tox = target, constrain = F, pointest="mean", method = 'rjags')

ptox = final_est$ndose[[1]]$est
ptoxL = final_est$ndose[[1]]$quantiles[1,]
ptoxU = final_est$ndose[[1]]$quantiles[5,]
mtd = which.min(abs(ptox - target))
if (plt) pp = plt_func(path, ptox, ptoxL, ptoxU, target, mtd, p_true_add, dose_add) else pp = NULL
return(list(path = path, ptox = ptox, ptoxL = ptoxL, ptoxU = ptoxU, mtd = mtd, flag = add_flag, plt = pp))
}

```


## test a single trial

```{r}
set.seed(1)
res = single_run_add_dose(N = 36, n_per_cohort = 3, target = 0.33, p_true = p_true_1_add, p.tox0 = p.tox0_add, dose = dose_add, train_frac = 1/3, plt = T)
res
```

## multiple runs - case 1

```{r}
set.seed(1)
niter = 500
res1_add = replicate(niter, single_run_add_dose(N = 36, n_per_cohort = 3, target = 0.33, p_true = p_true_1_add, p.tox0 = p.tox0_add, dose = dose_add, train_frac = 1/3, plt = FALSE))
# visit frequency
print("Percent of patients assigned to each dose over all trials")
vf1_add_half1 = sapply(res1_add[1,], function(t) {
summary = NULL
for (i in 1:length(dose_add)) summary = c(summary, sum(t[1:(N*train_frac), 'dose'] == i))
return(summary)
})
vf1_add_half2 = sapply(res1_add[1,], function(t) {
summary = NULL
for (i in 1:length(dose_add)) summary = c(summary, sum(t[(N*train_frac+1):N, 'dose'] == i))
return(summary)
})
rowSums(vf1_add_half1) / (niter * N * train_frac)
rowSums(vf1_add_half2) / (niter * N * (1-train_frac))
# mtd
print("Proportion of simulated trials recommending dose as MTD")
# table(unlist(res1_add[5,])) / niter
sapply(1:length(dose_add), function(t) mean(unlist(res1_add[5,]) == t))
table(unlist(res1_add[6,])) / niter
# True ptox
print("True p's")
p_true_1_add
```


```{r}
# [1] "Percent of patients assigned to each dose over all trials"
# [1] 0.0000 0.5985 0.0345 0.1545 0.0860 0.1265
# [1] 0.1570 0.1740 0.3145 0.2960 0.0550 0.0035
# [1] "Proportion of simulated trials recommending dose as MTD"
# [1] 0.056 0.282 0.464 0.190 0.008 0.000
# 
# FALSE  TRUE 
#  0.21  0.79 
# [1] "True p's"
# [1] 0.125 0.250 0.300 0.500 0.600 0.700
```


## multiple runs - case 2

```{r}
set.seed(1)
niter = 500
res2_add = replicate(niter, single_run_add_dose(N = 36, n_per_cohort = 3, target = 0.33, p_true = p_true_2_add, p.tox0 = p.tox0_add, dose = dose_add, train_frac = 1/3, plt = FALSE))
# visit frequency
print("Percent of patients assigned to each dose over all trials")
vf2_add_half1 = sapply(res2_add[1,], function(t) {
summary = NULL
for (i in 1:length(dose_add)) summary = c(summary, sum(t[1:(N*train_frac), 'dose'] == i))
return(summary)
})
vf2_add_half2 = sapply(res2_add[1,], function(t) {
summary = NULL
for (i in 1:length(dose_add)) summary = c(summary, sum(t[(N*train_frac+1):N, 'dose'] == i))
return(summary)
})
rowSums(vf2_add_half1) / (niter * N * train_frac)
rowSums(vf2_add_half2) / (niter * N * (1-train_frac))
# mtd
print("Proportion of simulated trials recommending dose as MTD")
# table(unlist(res1_add[5,])) / niter
sapply(1:length(dose_add), function(t) mean(unlist(res2_add[5,]) == t))
table(unlist(res2_add[6,])) / niter
# True ptox
print("True p's")
p_true_2_add
```


```{r}
# [1] "Percent of patients assigned to each dose over all trials"
# [1] 0.000000000 0.342666667 0.001333333 0.045333333 0.127333333 0.483333333
# [1] 0.018666667 0.003333333 0.016444444 0.117777778 0.558888889 0.284888889
# [1] "Proportion of simulated trials recommending dose as MTD"
# [1] 0.000 0.000 0.002 0.096 0.676 0.226
# 
# FALSE  TRUE 
# 0.834 0.166 
# [1] "True p's"
# [1] 0.005 0.010 0.050 0.200 0.300 0.500
```




