---
title: "Final Project: Bayesian Inference Clinical Trials and Nonparametric Models"
author: "Non-stat Team: Captain: Xinyi Zhang, Members: ..."
date: "`r Sys.Date()`"
output: pdf_document
---

# Introduction

- Background of the study

# Literature Review

- Review of the CRM design for phase I dose-finding trials

# Methodology

- Description of the logistic regression model used

## Model Reproduction

- Reproduce results in Figures 6 & 7

### Figure 6

```{r}
d = c(0.5, 1, 3, 5, 6)
y = c(0, 0, 0, 0, 0, 1, 0, 0, 1, 0, 0, 0, 0, 1, 0, 1, 0, 1)
x = c(rep(d[1], 3), rep(d[3], 3), rep(d[4], 12))
data = data.frame(x, y)
```

Fit logit model

```{r}
logit = glm(y ~ x, data = data, family = "binomial")
x_pred = with(data, data.frame(x = d))
tbl = cbind(x_pred, predict(logit, newdata = x_pred, type = "link", se = TRUE))
tbl = within(tbl, {p = plogis(fit); lb = plogis(fit - (1.96 * se.fit)); ub <- plogis(fit + (1.96 * se.fit))})
tbl
```

plot

```{r}
library(ggplot2)
ggplot(tbl) + 
geom_line(aes(x, p)) + geom_point(aes(x, p), shape = 21, colour = "black", fill = "white", size = 2, stroke = 1) + 
geom_line(aes(x, ub), linetype = 2) + geom_point(aes(x, ub), shape = 21, colour = "black", fill = "white", size = 2, stroke = 1) +
geom_line(aes(x, lb), linetype = 2) + geom_point(aes(x, lb), shape = 21, colour = "black", fill = "white", size = 2, stroke = 1) +
geom_hline(yintercept = 0.33, col = 'red', linetype = 2, linewidth = 0.8) + 
theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
scale_x_continuous(breaks = d) + labs(title = 'Repeat Figure 6(b)', x = 'Dose', y = 'Probability of DLT')
```

### Figure 7 

```{r}
d = c(10, 20, 40, 100, 200, 400, seq(800, 6400, by = 800))
y = rep(0, 37)
y[c(26, 35, 37)] = 1co
x = c(d[1:11], rep(d[10], 3), rep(d[11], 4), rep(d[12], 8), rep(d[13], 6), rep(d[14], 5))
data = data.frame(x, y)
```

Fit logit model

```{r}
logit = glm(y ~ x, data = data, family = "binomial")
x_pred = with(data, data.frame(x = d))
tbl = cbind(x_pred, predict(logit, newdata = x_pred, type = "link", se = TRUE))
tbl = within(tbl, {p = plogis(fit); lb = plogis(fit - (1.96 * se.fit)); ub <- plogis(fit + (1.96 * se.fit))})
tbl
```

plot

```{r}
ggplot(tbl) + 
geom_line(aes(x, p)) + geom_point(aes(x, p), shape = 21, colour = "black", fill = "white", size = 2, stroke = 1) + 
geom_line(aes(x, ub), linetype = 2) + geom_point(aes(x, ub), shape = 21, colour = "black", fill = "white", size = 2, stroke = 1) +
geom_line(aes(x, lb), linetype = 2) + geom_point(aes(x, lb), shape = 21, colour = "black", fill = "white", size = 2, stroke = 1) +
geom_hline(yintercept = 0.2, col = 'red', linetype = 2, linewidth = 0.8) + 
theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
scale_x_continuous(breaks = d[-(2:6)]) + labs(title = 'Repeat Figure 7(b)', x = 'Dose', y = 'Probability of DLT')
```


## Simulation Studies
```{r}
# R code for simulation studies goes here
```

## Results and Discussion
Answers to the questions

## Conclusion
Summary of the report

## Appendix
Any additional R code goes here






